---
title: "HarvardX::Data Science: Data Visualization:: exercise sample code"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# *Load package first*

``` {r package_loading}
library(dplyr)
library(ggplot2)
library(RColorBrewer) # special color pallette
library(dslabs)
library(plotly) # interactive plot
data(us_contagious_diseases)

```

# *Tile_plot*

 - The sample code given creates a tile plot showing the rate of smallpox cases per population. 

## Code


```{r smallpox_tile_s}
the_disease = "Smallpox"
dat <- us_contagious_diseases %>% 
   filter(!state%in%c("Hawaii","Alaska") & disease == the_disease & weeks_reporting >=10) %>% 
   mutate(rate = count / population * 10000) %>% 
   mutate(state = reorder(state, rate))

Smallpox_tile <- dat %>% ggplot(aes(year, state, fill = rate)) + 
  geom_tile(color = "grey50") + 
  scale_x_continuous(expand=c(0,0)) + 
  scale_fill_gradientn(colors = brewer.pal(9, "Reds"), trans = "sqrt") + 
  theme_minimal() + 
  theme(panel.grid = element_blank()) + 
  ggtitle(the_disease) + 
  ylab("") + 
  xlab("")
```

## Plotting

> Static plotting

```{r Smallpox_tile_s, echo=FALSE}
plot(Smallpox_tile)
```

> Interactive plotting

```{r Smallpox_tile_i, echo=FALSE}
ggplotly(Smallpox_tile)
```

# *Time series_plot*

- The sample code given creates a time series plot showing the rate of smallpox cases per population by state. 

## Code

> R code as follows:

```{r smallpox_time_series}
the_disease = "Smallpox"
dat <- us_contagious_diseases %>%
   filter(!state%in%c("Hawaii","Alaska") & disease == the_disease & weeks_reporting >= 10) %>%
   mutate(rate = count / population * 10000) %>%
   mutate(state = reorder(state, rate))

avg <- us_contagious_diseases %>%
  filter(disease==the_disease) %>% group_by(year) %>%
  summarize(us_rate = sum(count, na.rm=TRUE)/sum(population, na.rm=TRUE)*10000)

Smallpox_time_series <- dat %>% ggplot() +
  geom_line(aes(year, rate, group = state),  color = "grey50", 
            show.legend = FALSE, alpha = 0.2, size = 1) +
  geom_line(mapping = aes(year, us_rate),  data = avg, size = 1, color = "black") +
  scale_y_continuous(trans = "sqrt", breaks = c(5,25,125,300)) + 
  ggtitle("Cases per 10,000 by state") + 
  xlab("") + 
  ylab("") +
  geom_text(data = data.frame(x=1955, y=50), mapping = aes(x, y, label="US average"), color="black") + 
  geom_vline(xintercept=1960, col = "blue")
```

## Plotting

> Static plotting

```{r Smallpox_time_series_s, echo=FALSE}
plot(Smallpox_time_series)
```

> Interactive plotting

```{r Smallpox_time_series_i, echo=FALSE}
ggplotly(Smallpox_time_series)
```


> Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
